Critiquing the System Usability Scale from a Questionnaire Design Perspective: Beware of Acquiescence Bias

### 2013-09-16 paper_draft3.Rnw
phrasing updates from gdoc

With the increased focus on developing highly usable products, it has become highly important to actually measure the perceived usability of products or systems. Especially during usability studies, large or small, standardized questionnaires are being used widely for such measurement. Commonly used such questionnaires, in order of its first publication, include the Computer User Satisfaction Inventory (CUSI) \cite{kirakowski1987}, the Questionnaire for User Interface Satisfaction (QUIS) \cite{chin1988}, the After Scenario Questionnaire (ASQ) \cite{lewis1991}, the Software Usability Measurement Inventory (SUMI) \cite{kirakowski1993}, the Computer System Usability Questionnaires (CSUQ) \cite{lewis1995}, and the System Usability Scale (SUS) \cite{brooke1996}, among several others.

Out of all usability measurement questionnaires, the SUS has received the highest level of adoption in both industry and academia, with hundreds of references across publications alone. The SUS is comprised of ten statements measured on a 5-point agreement scale, yielding a single score summarizing the usability assessment of the evaluated system. Over the years, the SUS has been used across a variety of different systems, including hardware, software, websites, applications, and even non-electronic products. However, since the time of its inception in 1986, research regarding the design of valid and reliable questionnaires has advanced significantly, with several insights that now challenge some of the foundations on which the SUS was developed. Most notably are research insights related to acquiescence bias \cite{smith1967,krosnickPresser2010,sarisKrosnickShaeffer2010}, its relationship to satisficing \cite{krosnick1991,krosnick1999}, and the use of scales with optimal lengths \cite{krosnickFabrigar1997, groves2004}.

The remainder of this note outlines such advances in questionnaire design research relevant to the original SUS, review the SUS in the context of those, propose an alternative version to conform with these insights, and finally compare the original SUS to the proposed alternative. Note that this note's intention is not to reduce the number of statements asked about in the SUS (as attempted by others \cite{finstad2010,lewis2013}), nor the identification and hence elimination of biases that do not relate to acquiescence. Instead, the goal of this note is to critique the wording of the statements and response options of the SUS on theoretical grounds with empirical evidence and propose an alternative scale.

@@

Even though first published in 1996, the SUS was developed in 1986 by John Brooke while working at Digital Equipment Corporation (DEC) in the UK. It was used as a ``quick and dirty" scale to be administered after usability studies on electronic office systems, such as DEC's VT100, a text-based terminal system. The SUS measures attitudes and perceptions regarding the effectiveness, efficiency, and satisfaction with a system (in accordance with the measures of usability defined in ISO 9241-11). To measure a system's usability on these dimensions, the SUS is comprised of ten statements (see \ref{tab:items} for their exact wording) which the respondent is asked to rate individually. The SUS uses a Likert scale, established by Rensis Likert in 1932 \cite{likert1932}, which allows questionnaire respondents to specify their level of agreement or disagreement with each of the statements on a symmetric five-point agreement scale, ranging from ``Strongly disagree" to ``Strongly agree" (see Figure \ref{fig:originalScale}). Note that only its endpoints are labeled, while additionally all five scale items numbered from 1 to 5. As already noted in Brooke's initial work \cite{brooke1996}, the phrasing of the statement strongly influences the expressed level of agreement; hence, during the development of the SUS, statements that received the most extreme responses were selected. When analyzing responses to the SUS, the individual responses are consolidated into a single score to represent the global usability assessment for that system and to enable cross-system comparisons. To ensure that this summation is possible, all items of the SUS need to be evaluated by each respondent.

@@

One of the questionnaire biases that has been researched thoroughly is that of acquiescence bias, the insight from which are also highly relevant to the SUS. Acquiescence bias is the tendency of a respondent to be more likely to agree with a given statement independent of its substance \cite{smith1967}. Several aspects of questionnaire design contribute to the likelihood for acquiescence bias when responding to a questionnaire. First, when given a non-neutral statement, respondents are more likely to think of reasons why the statement is true, rather than expending cognitive effort to consider reasons for disagreement. This form of shortcutting the question answer process is often referred to as satisficing \cite{krosnick1991}. Second, respondents with lower self-perceived status assume the questionnaire administrator agrees with the posed statement, resulting in deferential agreement bias \cite{sarisKrosnickShaeffer2010}. Similarly, respondents whose personality naturally skews towards agreeableness are more likely to suffer from acquiescence bias \cite{sarisKrosnickShaeffer2010}. Finally, if the respondent's cognitive ability or motivation is lower, acquiescence bias is more likely \cite{krosnickNarayanSmith1996}.

Acquiescence bias is the strongest when presented with binary agree/disagree, yes/no, or true/false answer options \cite{smith1967}; however, similar effects have been shown for agreement scales (such as the Likert scale) \cite{sarisKrosnickShaeffer2010}. To minimize this bias, research has proposed to refer to the underlying construct in a neutral, non-leading question and offer a neutral scale, instead of using statements with agreement scales \cite{sarisKrosnickShaeffer2010}. Another suggestion to minimize this bias has been to use reverse-keyed constructs, i.e., the same construct is asked positively and negatively in the same survey, the raw scores of which are then combined to correct for acquiescence bias.

Picking the most valid and reliable rating scale for a given question has been another heavily researched topic, especially when moving away from agreement scales. First, research has shown that scale points that are fully labeled as compared to those that just use numbers optimize reliability and minimize bias \cite{groves2004}. Second, the scale length and its items depends on the nature of the construct being measured, i.e., if the construct is unipolar or bipolar in nature. Unipolar constructs range from zero to an extreme amount and are best measured on a 5-point scale, optimizing reliability while minimizing respondent burden \cite{krosnickFabrigar1997}. Bipolar constructs, on the other hand, range from an extreme negative to an extreme positive with a natural midpoint, which are best measured with a 7-point rating scale to maximize reliability and data differentiation \cite{krosnickFabrigar1997}.

\section{Acquiescence in the SUS}

In this section, we inspect the original SUS for weaknesses in regards to acquiescence bias, based on questionnaire design research insights mentioned in the previous section. It also describes an experimental evaluation to identify the effects of acquiescence bias in the SUS. The item numbers of the SUS referred to in the remainder of this note correspond to those in Table \ref{tab:items}.
\subsection{Heuristic review}

Each of the ten items of the SUS is constructed of a non-neutral statement and an agreement scale. In accordance with questionnaire design research \cite{smith1967}, this particular design encourages the effects of acquiescence bias, i.e., the likelihood for the respondents to simply agree with the statement provided. As an example, let's review item 3 of the SUS: ``I thought the system was easy to use." In this case, respondents are likely to expend unproportional effort on finding reasons that confirm that the system is actually easy to use, instead of thinking about aspects that would contribute to the system being interpreted as difficult to use. As a result, respondents are led towards a more agreeable response in line with ``easy to use" as expressed in the given statement. As another example, let's take a look at item 9 in the original SUS (``I felt very confident using the system"). This statement leads the respondent towards higher confidence, as that's what the statement suggests and since the agreement scale increases the likelihood to agree that the system actually made them feel confident. Additionally, the nature of the agreement scale leads respondents further towards agreement as disagreeing with anyone requires courage and cognitive effort \cite{sarisKrosnickShaeffer2010}. As the SUS is often administered at the end of usability studies, hence, not anonymously and with the questionnaire administrator present, this effect may be especially strong. As all other items of the SUS are phrased in the exact same way, this consideration applies equally across all of the SUS.

Notably, some of the items in the SUS have been reverse-keyed to ask about the same construct from opposite directions, as for example in items 3 (``easy to use") and 8 (``cumbersome to use"). This suggests that potential acquiescence bias effects may cancel each other out; however, this claim needs further evaluation. Nevertheless, this reverse-keyed approach is used only for a subset of the items in the SUS, hence, all other items (1, 2, 4, 5, 6, 7, 9, 10) continue to suffer from significant acquiescence biases overall. While some of the items (e.g., 4 and 10) are reasonably similar, they are still asking about a slightly different underlying constructs and cannot simply used to cancel out biases.

The SUS' agreement scale is offered as a bipolar scale with 5 items. As agreement/disgreement is a bipolar scale in its nature, the most appropriate scale length should be 7 points, to maximize reliability and data differentiation \cite{krosnickFabrigar1997}. However, when transforming the current agreement scale into construct-specific, neutral scales, the different scales and their lengths then depend on the nature of the underlying construct being measured, i.e., if the construct is unipolar or bipolar in nature \cite{krosnickFabrigar1997}.

\subsection{Experimental Evaluation of Acquiescence Bias}

\subsubsection{Experiment setup}
Participants of a massive open online course offered by Stanford University were asked to complete an optional post-course survey. The survey received 1746 responses. Respondents were randomly assigned to one of three weighted groups: 25\% were presented with the (original) SUS (n=439), 25\% with the reversed SUS (n=438), and 50\% received an example of a more robust scale proposed in this note (n=869) (see Table \ref{tab:items} for scale details). The system that respondents were asked to evaluate comprised of the course sites for browsing and watching lecture videos. The rest of the survey was the same for all respondents and contained typical course assessment questions.

@@

\begin{table}[h]
\small
\centering
\begin{tabular}{rccccr}
\toprule
 & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Reversed} \\
\cmidrule(r){2-3} \cmidrule(r){4-5}
Item \# & M & SD & M & SD & p value \\
\midrule
1 & 3.09 & 0.90 & 3.21 & 1 & 0.002\\
2 & 3.28 & 1.06 & 3.23 & 0.86 & 0.006\\
3 & 3.30 & 0.84 & 3.45 & 0.89 & \textless0.001\\
4 & 3.54 & 0.92 & 3.08 & 1.23 & \textless0.001\\
5 & 3.00 & 0.92 & 3.10 & 1.01 & 0.013\\
6 & 3.26 & 1.02 & 2.89 & 1.14 & \textless0.001\\
7 & 3.12 & 0.90 & 3.09 & 0.95 & 0.868\\
8 & 3.10 & 1.18 & 3.33 & 0.84 & 0.173\\
9 & 3.29 & 0.89 & 3.45 & 0.91 & \textless0.001\\
10 & 3.26 & 1.06 & 2.35 & 1.49 & \textless0.001\\
\midrule
overall & 80.58 & 16.14 & 77.92 & 14.23 & \textless0.001\\
\bottomrule
\end{tabular}
\caption{Original and reversed SUS scores providing strong evidence for acquiescence bias.}
\label{tab:acqui}
\end{table}
<<eval=FALSE>>=
testItems=function(i, oa, ra){
    print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(wilcox.test(oa,ra)$p.value,5), sep=" & "))
}
for(i in 1:10){
    testItems(i, m.orisus[,i], m.revsus[,i])
}
testItems("all", convSUS(m.orisus[,1:10]), convSUS(m.revsus[,1:10]))
@
\section{Alternative Proposal}

@@

\section{Acknowledgments}
We are grateful to the instructors of the two Stanford courses for allowing us to conduct survey experiments in their post-course surveys.

>>>>

With the increased focus on developing highly usable products, it has become highly important to actually measure the perceived usability of products or systems. Especially during usability studies, large or small, standardized questionnaires are being used widely for such measurement. Commonly used such questionnaires, include the Computer User Satisfaction Inventory (CUSI) \cite{kirakowski1987}, the Questionnaire for User Interface Satisfaction (QUIS) \cite{chin1988}, the After Scenario Questionnaire (ASQ) \cite{lewis1991}, the Computer System Usability Questionnaires (CSUQ) \cite{lewis1995}, and the System Usability Scale (SUS) \cite{brooke1996}, among several others.

Out of all usability measurement questionnaires, the SUS has received the highest level of adoption in both industry and academia, with hundreds of references across publications alone. The SUS is comprised of ten statements measured on a 5-point agreement scale, yielding a single score summarizing the usability assessment of the evaluated system. Over the years, the SUS has been used across a variety of different systems, including hardware, software, websites, applications, and even non-electronic products. However, since the time of its inception in 1986, research regarding the design of valid and reliable questionnaires has advanced significantly, with insights that now challenge some of the foundations of the SUS. Most notably are insights related to acquiescence bias \cite{smith1967,sarisKrosnickShaeffer2010}, satisficing \cite{krosnick1991,krosnick1999}, and scales lengths \cite{krosnickFabrigar1997, groves2004}.

The remainder of this note outlines such advances in questionnaire design research relevant to the SUS, reviews the original SUS in the context of those, proposes an alternative version to conform with these insights, and finally compares the SUS to the proposed alternative. Note that this note's intention is not to reduce the number of statements asked about in the SUS, nor the identification and hence elimination of biases that do not relate to acquiescence. Instead, the goal of this note is to critique the wording of the statements and response options of the SUS on theoretical grounds with empirical evidence and propose an alternative scale.

@@

Although first published in 1996, the SUS was developed in 1986 by John Brooke while working at Digital Equipment Corporation (DEC). It was used as a ``quick and dirty" scale administered after usability studies on electronic office systems, such as DEC's VT100, a text-based terminal system. The SUS measures attitudes regarding the effectiveness, efficiency, and satisfaction with a system (in accordance with the measures of usability defined in ISO 9241-11). The SUS is comprised of ten statements (see \ref{tab:items} for their exact wording) which the respondent is asked to rate individually. The SUS uses a Likert scale, established by Rensis Likert in 1932 \cite{likert1932}, for respondents to specify their level of agreement or disagreement with each of the statements on a symmetric 5-point agreement scale, ranging from ``Strongly disagree" to ``Strongly agree". Note that only its endpoints are labeled, while additionally all five scale items numbered from 1 to 5 (see Figure \ref{fig:originalScale}). As already noted in Brooke's initial work \cite{brooke1996}, the phrasing of the statements strongly influences the expressed level of agreement; hence, statements that received the most extreme responses were selected. When analyzing responses, the data are consolidated into a single score to represent the global usability assessment for that system and to enable cross-system comparisons. To ensure that this summation is possible, all items of the SUS need to be evaluated respondents.

@@

A questionnaire bias that is highly relevant to the SUS and has been researched extensively is acquiescence response bias: the tendency of a respondent to agree with a given statement independent of its substance \cite{smith1967}. Several aspects of questionnaire design contribute to the likelihood for acquiescence bias when responding to a questionnaire. First, when given a non-neutral statement, respondents are more likely to think of reasons why the statement is true, rather than affording cognitive effort to consider reasons for disagreement. This form of shortcutting the question answer process is often referred to as satisficing \cite{krosnick1991}. Second, respondents with lower self-perceived status assume the questionnaire administrator agrees with the posed statement, resulting in deferential agreement bias \cite{sarisKrosnickShaeffer2010}. Similarly, respondents whose personality naturally skews towards agreeableness are more likely to suffer from acquiescence bias \cite{sarisKrosnickShaeffer2010}. Finally, acquiescence bias is stronger in respondents with lower cognitive ability or motivation \cite{krosnickNarayanSmith1996}.

Acquiescence bias is the strongest when presented with binary agree/disagree, yes/no, or true/false answer options \cite{smith1967}; however, similar effects have been shown for agreement scales (such as the Likert scale) \cite{sarisKrosnickShaeffer2010}. To minimize this bias, research has proposed to refer to the underlying construct in a neutral, non-leading question and offer a neutral scale, instead of using statements with agreement scales \cite{sarisKrosnickShaeffer2010}.

How to pick the most valid and reliable rating scale for a given question has been subject to much research. First, research has shown that fully labeled scale points compared to those that just use numbers improve reliability and reduce bias \cite{groves2004}. Second, the choice of scale length and labels depends on the nature of the construct being measured, i.e., if it is unipolar or bipolar in nature. Unipolar constructs range from zero to an extreme amount and are best measured on a 5-point scale, while bipolar constructs range from an extreme negative to an extreme positive with a natural midpoint and are best measured on a 7-point scale. Following these guidelines improves a scale’s reliability and maximizes data differentiation while minimizing the cognitive burden on respondents \cite{krosnickFabrigar1997}.

\section{Acquiescence Bias in the SUS}

In this section, we inspect the original SUS for weaknesses in regards to acquiescence bias. We also describe an experimental evaluation to identify the effects of acquiescence bias in the SUS.

\subsection{Heuristic review}

Each of the ten items of the SUS is constructed of a non-neutral statement and an agreement scale. In accordance with questionnaire design research \cite{smith1967}, this particular design induces acquiescence, i.e., an increased likelihood for the respondents to simply agree with the statement provided. For instance, consider item 3 of the SUS (see Table \ref{tab:items}): ``I thought the system was easy to use." Respondents are likely to spend relatively more effort on finding reasons that confirm rather than contradict the statement that the system is easy to use. As a result, respondents tend to be more agreeable towards the ``easy to use" description in the statement. Consider item 9 from the SUS as another example: ``I felt very confident using the system". This loaded phrasing of this statement in combination with the agreement scale induces respondents to report higher levels of confidence than they would otherwise. Note also that this item confounds two dimensions by containing ``very", because disagreement with feeling very confident is not equivalent to feeling unconfident, but could theoretically indicate extreme confidence.  Additionally, the nature of the agreement scale leads respondents further towards agreement as disagreeing with anyone requires courage and cognitive effort \cite{sarisKrosnickShaeffer2010}. This effect may be especially strong, as the SUS is frequently administered at the end of usability studies without anonymity and in the presence of the questionnaire administrator. These issues apply to all SUS items as they are phrased in the exact same way.

Notably, some of the items in the SUS have been reverse-keyed to ask about the same construct from opposite directions, as for example in items 3 (``easy to use") and 8 (``cumbersome to use"). It may be that potential acquiescence bias effects cancel each other out; however, this claim requires empirical evaluation. Nevertheless, this reverse-keyed approach is used only for a subset of the items in the SUS, hence, all other items (1, 2, 4, 5, 6, 7, 9, 10) remain vulnerable to significant acquiescence bias. While some items (e.g., 4 and 10) are reasonably similar, they effectively tap into different underlying constructs and are thus unlikely mitigating any biases.

The SUS' agreement scale is offered as a bipolar scale with five items. As agreement/disagreement is clearly bipolar, the appropriate scale length would be seven points to maximize reliability and data differentiation \cite{krosnickFabrigar1997}. However, when transforming the current agreement scale into construct-specific, neutral scales, the different scales and their lengths depend on the nature of the measured underlying construct, i.e. whether the construct is unipolar or bipolar in nature \cite{krosnickFabrigar1997}.

To provide empirical support for the principal claim that the SUS is vulnerable to acquiescence bias in particular, we ran an online survey experiment with 1746 participants who evaluated an online learning system.

\subsubsection{Experiment Setup}

Participants of a massive open online course were asked to complete an optional post-course survey. Respondents were randomly assigned to one of three weighted groups: 25\% were presented with the (original) SUS (n=439), 25\% with the reversed SUS (n=438), and 50\% received an example of a more robust scale proposed in this note (n=869) (see Table \ref{tab:items} for scale details). The system that respondents were asked to evaluate comprised of the course sites for browsing and watching lecture videos. The rest of the survey was the same for all respondents and contained typical course assessment questions.

@@

\begin{table}[h]
\small
\centering
\begin{tabular}{rccccr}
\toprule
 & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Reversed} \\
\cmidrule(r){2-3} \cmidrule(r){4-5}
Item \# & M & SD & M & SD & p value \\
\midrule
1 & 3.09 & 0.90 & 3.21 & 1.00 & 0.002\\
2 & 3.28 & 1.06 & 3.23 & 0.86 & 0.006\\
3 & 3.30 & 0.84 & 3.45 & 0.89 & \textless0.001\\
4 & 3.54 & 0.92 & 3.08 & 1.23 & \textless0.001\\
5 & 3.00 & 0.92 & 3.10 & 1.01 & 0.013\\
6 & 3.26 & 1.02 & 2.89 & 1.14 & \textless0.001\\
7 & 3.12 & 0.90 & 3.09 & 0.95 & 0.868\\
8 & 3.10 & 1.18 & 3.33 & 0.84 & 0.173\\
9 & 3.29 & 0.89 & 3.45 & 0.91 & \textless0.001\\
10 & 3.26 & 1.06 & 2.35 & 1.49 & \textless0.001\\
\midrule
overall & 80.58 & 16.14 & 77.92 & 14.23 & \textless0.001\\
\bottomrule
\end{tabular}
\caption{Original and reversed SUS scores providing strong evidence for acquiescence bias.}
\label{tab:acqui}
\end{table}
<<eval=FALSE>>=
testItems=function(i, oa, ra){
    print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(wilcox.test(oa,ra)$p.value,5), sep=" & "))
}
for(i in 1:10){
    testItems(i, m.orisus[,i], m.revsus[,i])
}
testItems("all", convSUS(m.orisus[,1:10]), convSUS(m.revsus[,1:10]))
@
\section{Alternative Proposal}

@@

\section{Acknowledgments}
We are grateful to the instructors of the two online courses for allowing us to conduct survey experiments in their post-course surveys.
